% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  a4paper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage[]{Segoe Print}
\ifPDFTeX\else  
    % xetex/luatex font selection
    \setmainfont[]{Segoe UI}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Quick Guide to the UK Senior Manager Regime},
  pdfauthor={Owen Vallis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Quick Guide to the UK Senior Manager Regime}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\textbf{Briefing Paper: The UK's Senior Manager and
Certification Regime (SMCR)}}
\author{Owen Vallis}
\date{2025-01-06}

\begin{document}
\maketitle

\renewcommand*\contentsname{Contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}

\begin{tcolorbox}[enhanced jigsaw, coltitle=black, leftrule=.75mm, bottomrule=.15mm, toptitle=1mm, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, arc=.35mm, toprule=.15mm, colbacktitle=quarto-callout-warning-color!10!white, left=2mm, opacityback=0, breakable, titlerule=0mm, colback=white, bottomtitle=1mm, title=\textcolor{quarto-callout-warning-color}{\faExclamationTriangle}\hspace{0.5em}{WARNING - Disclaimer}, opacitybacktitle=0.6]

This guide is not official regulatory language and may not correctly or
completely reflect the requirements of the regulations discussed. The
author is not associated with any regulatory body.

\end{tcolorbox}

\subsubsection{Introduction}\label{introduction}

The Senior Manager and Certification Regime (SMCR) is a regulatory
framework established by the UK's Prudential Regulation Authority (PRA)
and Financial Conduct Authority (FCA) to enhance accountability and
governance within financial services firms. Introduced in 2016, the SMCR
replaced the Approved Persons Regime and aims to ensure that senior
managers are personally accountable for their decisions, promote a
culture of compliance, and strengthen consumer protection. This briefing
outlines the key components, rules, and compliance requirements under
the SMCR.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Key Objectives of the
SMCR}\label{key-objectives-of-the-smcr}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Strengthen Accountability}: Ensure senior managers take
  responsibility for their decisions and actions.
\item
  \textbf{Enhance Governance}: Improve oversight and management of risks
  within firms.
\item
  \textbf{Foster Ethical Culture}: Encourage firms to prioritize
  integrity, transparency, and fair treatment of customers.
\item
  \textbf{Improve Consumer Protection}: Minimize misconduct and improve
  confidence in the financial system.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Core Components of the
SMCR}\label{core-components-of-the-smcr}

The SMCR comprises three main components:

\paragraph{1. Senior Managers Regime
(SMR)}\label{senior-managers-regime-smr}

This applies to individuals in key leadership roles responsible for
overseeing critical functions within a firm. Key requirements include:

\begin{itemize}
\tightlist
\item
  \textbf{Senior Management Functions (SMFs)}: Specific roles such as
  Chief Executive, Chief Risk Officer, and Heads of Key Business Areas
  must be approved by the regulator.
\item
  \textbf{Statements of Responsibilities (SoRs)}: Senior managers must
  provide a clear and detailed document outlining their individual
  responsibilities.
\item
  \textbf{Duty of Responsibility}: Senior managers can be held
  personally accountable if they fail to take reasonable steps to
  prevent regulatory breaches within their area of responsibility.
\end{itemize}

\paragraph{2. Certification Regime}\label{certification-regime}

This applies to employees who are not senior managers but hold positions
that could significantly impact customers or the firm's integrity. Key
elements include:

\begin{itemize}
\tightlist
\item
  \textbf{Certification Requirement}: Firms must annually certify that
  these individuals are fit and proper to perform their roles.
\item
  \textbf{Significant Harm Functions (SHFs)}: Roles include
  client-facing advisors, proprietary traders, and individuals managing
  critical systems.
\end{itemize}

\paragraph{3. Conduct Rules}\label{conduct-rules}

The Conduct Rules apply to almost all employees within financial
services firms, providing a baseline for behavior and ethical standards.
They are divided into:

\begin{itemize}
\tightlist
\item
  \textbf{Individual Conduct Rules}: Applicable to all staff, including
  requirements to act with integrity, provide appropriate care, and
  comply with regulations.
\item
  \textbf{Senior Manager Conduct Rules}: Additional rules for senior
  managers, emphasizing their responsibility for effective control and
  delegation.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Approved Functions under the
SMCR}\label{approved-functions-under-the-smcr}

The following is the full list of Senior Management Functions (SMFs)
requiring PRA or FCA approval:

\begin{longtable}[]{@{}lr@{}}
\toprule\noalign{}
Function & Identifier \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Chief Executive Officer & SMF1 \\
Chief Finance Officer & SMF2 \\
Executive Director & SMF3 \\
Chief Risk Officer & SMF4 \\
Head of Internal Audit & SMF5 \\
Head of Key Business Area & SMF6 \\
Group Entity Senior Manager Function & SMF7 \\
Chairman & SMF9 \\
Chair of the Risk Committee & SMF10 \\
Chair of the Audit Committee & SMF11 \\
Chair of the Remuneration Committee & SMF12 \\
Senior Independent Director & SMF14 \\
Compliance Oversight Function & SMF16 \\
Money Laundering Reporting & SMF17 \\
Other Overall Responsibility Function & SMF18 \\
\end{longtable}

\subsubsection{Certified Functions under the
SMCR}\label{certified-functions-under-the-smcr}

Certified functions are roles that do not require regulatory approval
but are deemed critical to a firm's integrity. These include:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Significant Management Functions}: Individuals managing large
  teams or key business units.
\item
  \textbf{Proprietary Traders}: Staff involved in trading activities
  that could pose significant risks to the firm.
\item
  \textbf{CASS Oversight Function}: Roles responsible for overseeing
  compliance with the Client Assets Sourcebook.
\item
  \textbf{Material Risk Takers}: Employees whose decisions could
  materially affect the firm's risk profile.
\item
  \textbf{Client Dealing Functions}: Roles involving advising,
  arranging, or dealing with customers.
\item
  \textbf{Algorithmic Trading Functions}: Individuals responsible for
  the design and oversight of algorithmic trading systems.
\end{enumerate}

To qualify as a certified function, the role must meet the following
criteria:

\begin{itemize}
\tightlist
\item
  The individual's actions or decisions could significantly impact the
  firm's business or customers.
\item
  The role requires specific skills, knowledge, or experience to ensure
  effective performance.
\item
  The individual must be certified annually by the firm as fit and
  proper to perform the function.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Conduct Rules}\label{conduct-rules-1}

\paragraph{Individual Conduct Rules}\label{individual-conduct-rules}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Integrity}: Act with honesty and integrity in all actions.
\item
  \textbf{Due Care, Skill, and Diligence}: Perform duties with care and
  attention to detail.
\item
  \textbf{Cooperation with Regulators}: Be open and cooperative with the
  FCA, PRA, and other regulatory bodies.
\item
  \textbf{Customer Interests}: Always act in the best interests of
  customers.
\item
  \textbf{Market Conduct}: Observe proper standards of market conduct in
  all activities.
\end{enumerate}

\paragraph{Senior Manager Conduct
Rules}\label{senior-manager-conduct-rules}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Reasonable Steps}: Take reasonable steps to ensure the firm's
  business is controlled effectively.
\item
  \textbf{Compliance}: Ensure the business complies with relevant
  regulatory requirements.
\item
  \textbf{Delegation Oversight}: Ensure any delegation of
  responsibilities is appropriate and effective.
\item
  \textbf{Disclosure}: Disclose appropriately any information of which
  the FCA or PRA would reasonably expect notice.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Compliance Requirements}\label{compliance-requirements}

\paragraph{1. Implementation and
Oversight}\label{implementation-and-oversight}

Firms must establish clear governance structures to ensure SMCR
compliance. This includes: - Identifying all Senior Management Functions
and Significant Harm Functions. - Preparing Statements of
Responsibilities for senior managers. - Conducting regular reviews of
roles and responsibilities.

\paragraph{2. Fitness and Propriety
Assessments}\label{fitness-and-propriety-assessments}

Firms must assess the suitability of senior managers and certified
individuals, focusing on: - \textbf{Honesty, Integrity, and Reputation}:
Ensuring individuals have a history of ethical behavior. -
\textbf{Competence and Capability}: Evaluating skills and experience
relevant to their role. - \textbf{Financial Soundness}: Checking for
financial stability and solvency.

\paragraph{3. Training and Awareness}\label{training-and-awareness}

Firms must provide regular training to employees on the Conduct Rules
and their implications. This ensures all staff understand their
responsibilities under the regime.

\paragraph{4. Recordkeeping}\label{recordkeeping}

Firms must maintain comprehensive records of: - Role allocations and
responsibilities. - Certification decisions and supporting evidence. -
Breaches of Conduct Rules and remedial actions taken.

\paragraph{5. Regulatory Reporting}\label{regulatory-reporting}

Firms are required to report significant changes, breaches, or
misconduct involving senior managers and certified individuals to the
FCA or PRA promptly.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Enforcement and
Penalties}\label{enforcement-and-penalties}

The SMCR introduces significant consequences for non-compliance,
including: - \textbf{Regulatory Action Against Individuals}: Senior
managers can face fines, bans, or criminal charges for breaches of their
Duty of Responsibility. - \textbf{Firm-Level Sanctions}: Firms may face
financial penalties, operational restrictions, or reputational damage
for systemic governance failures.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Challenges and Best Practices for
Firms}\label{challenges-and-best-practices-for-firms}

\textbf{Challenges}: 1. Ensuring clarity in the allocation of
responsibilities to avoid overlaps or gaps.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\item
  Maintaining robust systems for annual certification and ongoing
  monitoring.
\item
  Navigating regulatory scrutiny and expectations around governance.
\end{enumerate}

\textbf{Best Practices}:

\begin{itemize}
\tightlist
\item
  \textbf{Comprehensive Governance Frameworks}: Establish clear policies
  and procedures to manage responsibilities and compliance.
\item
  \textbf{Proactive Risk Management}: Regularly review and update risk
  management processes.
\item
  \textbf{Cultural Alignment}: Foster a culture of accountability and
  transparency across all levels of the organization.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{Conclusion}\label{conclusion}

The implementation of the SMCR represented a pivotal shift in the
regulation of financial services, emphasizing accountability,
governance, and ethical conduct. By adhering to its requirements, firms
not only mitigate regulatory risks but also strengthen their operational
resilience and consumer trust. With the regulatory landscape
continuously evolving, proactive engagement with the SMCR's principles
is essential for maintaining competitive and compliant operations in the
UK's financial sector.

\begin{tcolorbox}[enhanced jigsaw, coltitle=black, leftrule=.75mm, bottomrule=.15mm, toptitle=1mm, colframe=quarto-callout-caution-color-frame, rightrule=.15mm, arc=.35mm, toprule=.15mm, colbacktitle=quarto-callout-caution-color!10!white, left=2mm, opacityback=0, breakable, titlerule=0mm, colback=white, bottomtitle=1mm, title=\textcolor{quarto-callout-caution-color}{\faFire}\hspace{0.5em}{AI use}, opacitybacktitle=0.6]

An Ai tool was utilised in the drafting process for this guide.

\end{tcolorbox}




\end{document}
